### 2025-06-12 conteudo/methods/methods.tex
add more numerical results from students

This section outlines our approach to mitigating code duplication in the Linux kernel 
and describes the ethnographic methodology for engaging with the Linux community. 
Firstly, we present the participant observation research method explored by the author. 
Then, we present the methodology of the experiment done with the students.

@@

Given the tool evaluations demonstrating the tool's capabilities of finding duplications, 
we proceeded to understand if the duplications found can be mitigated and the kernel 
subsystem's maintainers' opinions about the duplications and their respective mitigations. 
To accomplish this objective, we initially proposed a participant-observation experiment, 
where we, people who have never contributed to the Linux kernel, act as contributors to 
the AMD Display driver and collect artifacts of our experience in the process.

As a first step of this experiment, we executed the tool in the AMD Display driver codebase. 
We manually analyzed the biggest duplications regarding the number of lines and raised them 
to two duplicated function pairs that we judged interesting to try to mitigate. We have 
chosen to approach the biggest duplications as proof of concept so that we can make 
significant impacts. The two duplicated function pairs chosen are:

@@

\subsection{Study with University Students}

@@

In the second stage of the Free Software Development course, we propose that students contribute 
directly to the ArKanjo tool development without help from the tool creator to verify whether the 
tool can be free software to which multiple people can contribute. The students approached the tool’s 
issue, and to our knowledge, they pointed to us in the survey feedback. Three students forming two 
groups worked in direct contributions, which was not a very expressive number compared with the 
initial experiment.

>>>>

This section outlines our approach to mitigating code duplication in the Linux kernel 
and describes the ethnographic methodology for engaging with the Linux community. 
Firstly, we present the participant observation research method explored by the author. 
Then, we present the methodology of the non-participant observation analyze of patches 
sent by the students.

@@

Given the tool evaluations demonstrating the tool's capabilities of finding duplications, 
we proceeded to understand if the duplications found can be mitigated and the kernel 
subsystem's maintainers' opinions about the duplications and their respective mitigations. 
To accomplish this objective, we proposed a participant-observation experiment, 
where we, people who have never contributed to the Linux kernel, act as contributors to 
the AMD Display driver and collect artifacts of our experience in the process.

We executed the tool in the AMD Display driver codebase. 
We manually analyzed the biggest duplications regarding the number of lines and raised them 
to two duplicated function pairs that we judged interesting to try to mitigate. We have 
chosen to approach the biggest duplications as proof of concept so that we can make 
significant impacts. The two duplicated function pairs chosen are:

@@

\subsection{Non-participant Observation Study with University Students}

@@

In the second stage of the Free Software Development course, we propose that students contribute 
directly to the ArKanjo tool development without help from the tool creator to verify whether the 
tool can be free software to which multiple people can contribute. The students approached the tool’s 
issue, and to our knowledge, they pointed to us in the survey feedback. Three students forming two 
groups worked in direct contributions, which was not a very expressive number compared with the 
initial stage of the course.

### 2025-06-12 conteudo/results/participant.tex
add more numerical results from students

For our initial participant observation, we attempted to mitigate two code 
duplications found by the ArKanjo tool in the AMD Display driver by applying 
our systematic approach presented in \ref{subsubsec:systematic}. 
As presented in section \ref{sec:meteth},
we opted to apply the systematic approach in the functions
\textit{offset\_to\_id} and \textit{phy\_id\_to\_atom} in the folders 
\textit{dc/gpio} and \text{dc/bios}, respectively.

>>>>

For our participant observation, we attempted to mitigate two code 
duplications found by the ArKanjo tool in the AMD Display driver by applying 
our systematic approach presented in \ref{subsubsec:systematic}. 
As presented in section \ref{sec:meteth},
we opted to apply the systematic approach in the functions
\textit{offset\_to\_id} and \textit{phy\_id\_to\_atom} in the folders 
\textit{dc/gpio} and \text{dc/bios}, respectively.

### 2025-06-12 conteudo/results/results.tex
add more numerical results from students

In this chapter, we present and analyze the results of the experiments and methods 
used in our research. Following the chronology of the study, we start by presenting the 
evaluation of the ArKanjo tool through both a method from the literature and an empirical 
analysis. Then, we present and discuss the experience of the initial participant 
observation methodology, where we mitigate a code duplication on the AMD Display Driver 
and send it as a patch. Finally, we show and discuss our discoveries on the experiment 
done with students at the University of São Paulo, where we propose to students to mitigate 
duplications found by the ArKanjo tool at IIO subsystem and the AMD Display Driver, and 
document their founds, opinions, and experiences.

>>>>

In this chapter, we present and analyze the results of the experiments and methods 
used in our research. Following the chronology of the study, we start by presenting the 
evaluation of the ArKanjo tool through both a method from the literature and an empirical 
analysis. Then, we present and discuss the experience of the initial participant 
observation methodology, where we mitigate a code duplication on the AMD Display Driver 
and send it as a patch. Finally, we show and discuss our discoveries of the non-participant observation
of the students approaches to mitigate duplications and patches's discussions.

### 2025-06-12 conteudo/results/students.tex
add more numerical results from students

As presented in Chapter \ref{cha:method}, we propose that students mitigate duplications 
found by the ArKanjo tool, send patches to the Linux subsystem maintainers, and document 
their experiences and learnings.

For the students who worked in the IIO subsystem, we gave them duplicated function pairs 
found by the tool, and their responsibility was to think of how to mitigate the duplication 
and send the patch to the maintainers. The student that worked in the AMD Display driver 
have the additional work to use the tool to find the duplications.

@@

Table \ref{tab:stu} summarizes students' groups' mitigation approaches regarding lines 
changed and refactor methods used. Seven groups applied the parameterized refactor method, 
and two applied the extractor method to their approaches to mitigate the duplications. 
The other three groups do not use any refactor methods presented in the literature.

>>>>

As presented in Chapter \ref{cha:method}. It was proposed to students, as one of the activities 
of the Free Software course, that they make the first-time contribution to the Linux kernel to 
mitigate duplication found by the ArKanjo tool while sending patches to the Linux subsystem 
maintainers and documenting their experiences and learnings.

For the students who worked in the IIO subsystem, the teaching assistant gave them duplicated 
function pairs 
found by the tool, and their responsibility was to think of how to mitigate the duplication 
and send the patch to the maintainers. The student that worked in the AMD Display driver 
have the additional work to use the tool to find the duplications.

@@

Table \ref{tab:stu} summarizes students' groups' mitigation approaches. Seven groups applied 
the parameterized refactor method, 
and two applied the extractor method to their approaches to mitigate the duplications. 
The other three groups do not use any refactor methods presented in the literature.  
Five groups got their patches accepted and merged in Linux; four groups got their patches rejected, 
and one group is still working on the patch submission, being in the fifth version of the patch at 
the time this document was written. Two groups decided not to continue sending the patches after 
they made initial versions.

### 2025-06-12 conteudo/results/table_students.tex
add more numerical results from students

\begin{table}
\begin{tabular}{ | c | c | c | c | m{6em} | }

@@

\textbf{Group} & \textbf{Subsystem} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
\\ \hline

1 & IIO & 7 & 14 & None \\ \hline
2 & IIO & 23 & 24 & Parameterize method \\ \hline
3 & IIO & 12 & 30 & Parameterize method \\ \hline
4 & IIO & 2 & 7 & None \\ \hline
5 & IIO & 17 & 70 & Parameterize method \\ \hline
6 & IIO & - & - & Parameterize method \\ \hline
7 & IIO & 21 & 22  & None \\ \hline
8 & IIO & - & - & None \\ \hline
9 & IIO & 33 & 52 & Parameterize method \\ \hline
10 & IIO & 32 & 58 & Parameterize method \\ \hline
11 & IIO & - & - & Parameterize method \\ \hline
12 & AMD & 89 & 489 & None \\ \hline

>>>>

\begin{table}
\begin{tabular}{ | c | c | c | m{6em} | }

@@

\textbf{Group} & \textbf{Subsystem} & \textbf{Patch Status} & \textbf{Refactoring methods used} 
\\ \hline

1 & IIO & Rejected & None \\ \hline
2 & IIO & Rejected & Parameterize method \\ \hline
3 & IIO & Accepted & Parameterize method \\ \hline
4 & IIO & Accepted & None \\ \hline
5 & IIO & Accepted  & Parameterize method \\ \hline
6 & IIO & Students stopped working on it & Parameterize method \\ \hline
7 & IIO & Accepted  & Extract method \\ \hline
8 & IIO & Rejected & Extract method \\ \hline
9 & IIO & Work in progress & Parameterize method \\ \hline
10 & IIO & Rejected & Parameterize method \\ \hline
11 & IIO & Students stopped working on it & Parameterize method \\ \hline
12 & AMD & Accepted & None \\ \hline
